The Next Economic Era

### 2014-02-19 draft1.tex
began incentives section 1.00

			The third point will arguably prove to be the most important to our thesis: the fact that the free trade of property involves two sides, production and consumption. It is so important that it will probably 
			take us a couple chapters to truly develop a deep understanding of the issue. Nevertheless, it is worth having a first glance. A key feature of a capitalist system is that it incentives people, to a certain 
			extent, based on greed. This notion of "greed" can be quite innocent. It involves our needs, such as food and hopefully shelter, and our wants, which can be anything from a cup of tea to less modest pursuits 
			like a private jet. The useful aspect of this desire for things others produce is that it is generally widespread, and sufficiently motivating for people to work. Effort and reward certainly do not always go 
			hand in hand linearly, and so it is not a simple relationship. But overall, in such a system if you want things others produce, you'll have to produce yourself. I argue that the subtleties of this create some of 
			the most profound problems with capitalism today. It will take me several chapters to attempt to elaborate on this, but I hope that will allow me to convey the main ideas with clarity. Before we do that, let's 
			have a slightly deeper look at our incentives to work and consume, since they are so central to a capitalist system, and will in all likelihood be so for a new economic system.

		% What characteristics a realistic incentive structure must have
		\section{Incentives}
			% i.e. now that we know a bit more about capitalism, reasons it used to and once again has superseded central planning

		 % The problem of wealth distribution, with a number of illustrated examples
		 \section{A Rich Man's Game}

		% The problem of job dissatisfaction, with...
		\section{My Job Sucks}

		% The Economic case against inequality and disillusionment (vs the moral case...)
		% e.g. skew to markets for the wealthy only, diminishing marginal returns...
		\section{All Work and No Play Makes Jack a Broken Economy}

>>>>

			The third point will arguably prove to be the most important to our thesis: the fact that the free trade of property involves two sides, production and consumption. It is so important that it will probably 
			take us a couple chapters to truly develop a deep understanding of the issue. Nevertheless, it is worth having a first glance. A key feature of a capitalist system is that it incentives people, to a certain 
			extent, based on greed. This notion of "greed" can be quite innocent. It involves our needs, such as food and hopefully shelter, and our wants, which can be anything from a cup of tea to less modest pursuits 
			like a private jet. The useful aspect of this desire for things others produce is that it is generally widespread, and sufficiently motivating for people to work. Effort and reward certainly do not always go 
			hand in hand linearly, and so it is not a simple relationship. But overall, in such a system if you want things others produce, you'll have to produce yourself. I argue that the subtleties of this create some of 
			the most profound problems with capitalism today. It will take me several chapters to attempt to elaborate on this, but I hope that will allow me to convey the main ideas with clarity. Before we do that, let's 
			have a slightly deeper look at our incentives to work and consume, since they are so central to a capitalist system, and will in all likelihood be so for an improved economic system.

		% What characteristics a realistic incentive structure must have
		\section{Incentives}
			% i.e. now that we know a bit more about capitalism, reasons it used to and once again has superseded central planning
			% Intro: not psychology. We're going to look into incentives for people to work in a focused context.
			We now have an idea of the key components of a capitalist system. Why do we care about analyzing incentives in such a system? First of all, because there is a lot of merit 
			to capitalist incentives. Despite the drawbacks, people today are generally very motivated to work, often at very displeasing but economically productive jobs. Secondly, if we are going to 
			criticize the current system we will need to know which aspects of it already work well, which work less well, and whether or how we can fix the latter. The alternative is that in exploring the flaws 
			we propose extreme solutions that in fact worsen the illness. Later in this section I'll advance the idea that communism could have in part been such an extreme reaction to the problem of inequality endemic 
			to industrial capitalism. Indeed, there are many flavors of capitalism, and the vanilla of such flavors could very well be rooted in the first and second industrial revolutions. Let's have a look at how those 
			revolutions might have had an impact on our economic incentives. I'd like to stress that we'll be looking at incentives in the very narrow economic sense, as this is not a work in psychology. Regardless, that should 
			suffice in reaching some potent conclusions.\

			One could look at the causes of the industrial revolutions at multiple levels of granularity and across multiple dimensions. For example, one could look at the societal factors behind the free thinking 
			that enabled scientific inventions like the steam engine, or printing press before that. One could go a few hundred years back in time to trace the origins of modern practices like double-entry 
			bookkeeping by merchants in 14th century Venice \cite{VeniceMerchants}. Whatever the case, the First Industrial Revolution is generally taken to mean the massive changes that tools place
			across all forms of economic life from around the 1760s to the 1840s, and the Second Industrial Revolution from then up till the 1870s. I personally prefer the view that these changes 
			in fact represented part of a much broader and gradual evolution mainly in Western societies. As Acemoglu and Robinson have argued \cite{WhyNationsFail}, events like the Glorious Revolution of 1688
			in Britain, or the 1789 French Revolution, were part of an ongoing trend towards more and more inclusive institutions in certain parts of the world. These inclusive institutions essentially allowed a broader 
			segment of society to take part in economic and political life. Monarchs' powers were diluted, elected representative bodies were formed, and the rule of law began to apply more equally to all citizens.\

			This actually brings us to an extremely important point: the bidirectional interplay between incentives and capitalism. On one hand, incentives in no small part led to modern industrial capitalism. On the 
			other hand, this economic revolution has had all sorts of consequences on our economic incentives. With regards to the first direction of causality, although intellectual vibrancy no doubt had its role in 
			inventions like the steam engine, it was inclusive institutions like  well-defined property rights that allowed inventors to benefit economically from their discoveries, thereby massively 
			catalyzing scientific progress and economic productivity. In return, these large increases in economic productivity have led to significant changes in the way we organize ourselves as a 
			society, and in turn completely changed the dynamics of our incentives to work and consume.\

			% What were the 1st and 2nd Industrial Revolution
			% ## By how much exactly did output increase: which countries most, how much per capita in those countries vs in world
			% When did Communism/Socialism arise? What have people argued as its main causes?
			% Where does output per capita stand today, around the world?
			% -------
			% With that background, realities of what incentives we need. The market for labor: (try to bring in some stats)
			%	- how many people love their work, some surveys? how many enjoy...how many bear with it for long term purpose, how many just hate it..... ==> come up with a novel illustration/graph of this
			%	- if so many unfulfilled etc (I suspect), why work? ==> wage distribution ==> how much for needs, how much for wants, and so what are driving factors
			%	- what kind of jobs do people take up: SECTOR composition, INDUSTRY composition, TYPE composition (entrepreneur, sme, big biz) ==> bargaining power
			%		==> do look closely at bargaining power: suppliers and demanders of labor
			% segway to next section. Hint at idea of productive wells and who how the production is distributed. Spend some time emphasizing this as a MAJOR point of book

		 % The problem of wealth distribution, with a number of illustrated examples
		 % Now getting deep into point of wells of production and where the wealth goes
		 \section{A Rich Man's Game}
		 	% wealth distribution around the world
			% types of jobs and relative wealth: where is the money going

		% The problem of job dissatisfaction, with...
		\section{My Job Sucks}
			% Intro: this section is more fuzzy. More like a rally / oration on soul wrenching jobs. Will have to be inspirational. We are humans not machines, so a holistic book will include this part.

		% The Economic case against inequality and disillusionment (vs the moral case...)
		% e.g. skew to markets for the wealthy only, diminishing marginal returns...
		\section{All Work and No Play Makes Jack a Broken Economy}
			% yep worth writing about, especially after last more emotional section
